Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2023-05-29 2_SSL_NE/article/article_methods.tex

\begin{document}
Banana
\end{document}

>>>>

\begin{document}

\section{Methods and Data Collection}
\label{sec:methods}

The evolution of LED device architecture and performance as well as the progress in understanding the underlying physical phenomena are well covered in the scholarly literature and patents. However, information provided in such sources is insufficient for our goals on at least three accounts: First, existing work focuses only on selected performance parameters or overall device efficiency, rather than on providing a comprehensive coverage of the whole device sub-efficiencies for a particular LED product or design. Scientific publications also do not always disclose the underlying device architecture or the features responsible for the gains in performance. Second, not all relevant innovations are patented \cite{Pakes_1980,Fontana_2013}. In the case of LED patents in particular, our interviews with industry experts suggest that the propensity to patent is the highest for knowledge related to macroscopic device architecture and chemical composition of phosphors, and the lowest for knowledge related to manufacturing process improvements and microscopic chip architecture that is difficult to reconstruct by reverse engineering. This means that relying only on patent literature would bias results by unduly emphasizing some focus areas and de-emphasizing others. Third, scientific publications and patents typically focus on experimental devices, rather than commercial products. While new LED features, designs and manufacturing methods reported in these sources can potentially result in significant performance gains or cost reductions, it is difficult to ascertain if these improvements have since been adopted in industry.  Furthermore, information on LED manufacturing cost and the effect of process improvements on the total cost is highly proprietary. Estimates are occasionally reported in the scientific literature and company publications, but these often do not disclose which parts of the manufacturing process are responsible for the largest contribution to the overall cost, or which improvements led to cost reductions.

To overcome the limitations of these different methods for understanding technological progress, in this study we rely on a multi-method approach to data collection and analysis, the details of which are provided in Section 2 of the Supplementary Information document. Specifically, we combine information obtained from a systematic review of the primary scientific literature, device datasheets, relevant patents, and industry publications (SI Section 2.1) with information gained from semi-structured interviews with experts from academia and industry (SI Section 2.2), bottom-up manufacturing cost modelling (SI Section 2.3), and our own computations of device sub-efficiencies (SI Section 2.4). We then use this information to track the historical progress in white LED technology over time across the three groups of metrics identified above in the 'Metrics' section and identify its sources in innovation and technology spillovers.

\end{document}

### 2023-05-29 2_SSL_NE/article/article_supplementary.tex

\documentclass[10pt]{article}
% page setup
\usepackage[a4paper, total={6in, 8.75in}]{geometry}
\usepackage{parskip}
% formatting
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc} % use 8-bit T1 fonts
% cross-referencing
\usepackage{hyperref}
\usepackage{url}
\usepackage{doi}
% tables
\usepackage{booktabs}
% fonts
\usepackage{amsfonts}
\usepackage{microtype}
% s
\usepackage{graphicx}
\usepackage{float}
% math
\usepackage{amsmath}
% tables
\usepackage{tabularx}
\usepackage{colortbl}
% boxes
\usepackage{fancybox}
% tikz
\usepackage{tikz}
\usetikzlibrary{backgrounds}
\usetikzlibrary{arrows}
\usetikzlibrary{shapes,shapes.geometric,shapes.misc}
% landscape pages
\usepackage{pdflscape}
% chemical formulas
\usepackage{chemformula}
% smart references
\usepackage[capitalise, nameinlink]{cleveref}

\renewcommand{\thefigure}{SI\arabic{figure}}
\renewcommand{\thetable}{SI\arabic{table}}

@@

% this style is applied by default to any tikzpicture included via \tikzfig
\tikzstyle{tikzfig}=[baseline=-0.25em,scale=0.5]

% these are dummy properties used by TikZiT, but ignored by LaTex
\pgfkeys{/tikz/tikzit fill/.initial=0}
\pgfkeys{/tikz/tikzit draw/.initial=0}
\pgfkeys{/tikz/tikzit shape/.initial=0}
\pgfkeys{/tikz/tikzit category/.initial=0}

% standard layers used in .tikz files
\pgfdeclarelayer{edgelayer}
\pgfdeclarelayer{nodelayer}
\pgfsetlayers{background,edgelayer,nodelayer,main}

% style for blank nodes
\tikzstyle{none}=[inner sep=0mm]

% include a .tikz file
\newcommand{\tikzfig}[1]{%
{\tikzstyle{every picture}=[tikzfig]
\IfFileExists{#1.tikz}
  {\input{#1.tikz}}
  {%
    \IfFileExists{./figures/#1.tikz}
      {\input{./figures/#1.tikz}}
      {\tikz[baseline=-0.5em]{\node[draw=red,font=\color{red},fill=red!10!white] {\textit{#1}};}}%
  }}%
}

% the same as \tikzfig, but in a {center} environment
\newcommand{\ctikzfig}[1]{%
\begin{center}\rm
  \tikzfig{#1}
\end{center}}

\setcounter{secnumdepth}{4}

% fix strange self-loops, which are PGF/TikZ default
\tikzstyle{every loop}=[]
\input{./auxiliary/tikz_style_files/main.tikzstyles}

%%% custom definitions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\definecolor{silver}{rgb}{0.75, 0.75, 0.75}

%%% document metadata %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers  }
\date{February 2023}

\hypersetup{
    pdftitle={Supplementary Information},
    pdfauthor={Michael Weinold, Sergey Kolesnikov, Laura Diaz Anadon}
}

%%% document body %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{figure}[]
	\centering
    \includegraphics[width=\textwidth]{2_SSL_EST/article/figures/costmodel_sensitivity.pdf}
	\caption{Sensitivity analysis for selected cost model parameters. Panel a represents the tested range of variation of parameters indicated on horizontal axis along with their respective measurement units, as represented in \cref{tab:sensitivity}. Panel b shows the resulting range of variation in the manufacturing cost. Ranges are represented for three years considered in the cost model, where corresponding data is available, by whiskers of the following colour: 2003 – blue, 2012 – green, 2020 – red. Note the overall trend of decreasing sensitivity to cost model parameters as the number of die per wafer increases over time. Abbreviations: resp. - respective; yield sav. from insp. - yield savings from inspection.}
	\label{fig:sensitivity}
\end{figure}

@@

\begin{table}[]
\small
    \caption{Cost model sensitivity analysis parameter list. The results of the sensitivity analysis for the parameters in this table are presented in \cref{fig:sensitivity}.}
    \vspace{5mm}
    \begin{tabularx}{\textwidth}{ |X|X|l|l|l|l|l|l|X|}
        \hline
            \textit{Parameter} & \textit{Unit} & \textit{2003} & $\pm [\%]$ & \textit{2012} & $\pm [\%]$ & \textit{2020} & $\pm [\%]$ & Source \\
        \hline
            Cleanroom Cost & \text{USD}/m$^2$ & 3000 & +16,-16 & 3000 & +16,-16 & 3000 & +16,-16 & \cite{mddi1997cleanroom}\cite{ledcomv2} \newline \cite{bakshi2009euv}\cite{gajera2006process} \\
        \hline
            Person-hour & FTE & 100\% & +50,-50 & 100\% & +50,-50 & 100\% & +50,-50 & I \\
        \hline
            Equip. Discount & \% of \text{USD} & 0\% & +25,-25 & 0\% & +25,-25 & 0\% & +25,-25 & I \newline \cite{Appleyard_2001} \\
        \hline
            Overall Yield & \% & 100\% & +25,-25 & 100\% & +25,-15 & 100\% & +25,-25 & I, \cite{lumi2012yield}\cite{ledsmag2012} \newline \cite{systemplus2015reverse}\cite{ledcomv2} \\
        \hline
            Inspec. Yield Savings & \%/inspec. & 0.5\% & +80,-80 & 0.5\% & +80,-80 & 0.5\% & +80,-80 & \cite{mckinseyyield} \\
        \hline
            Overall Throughput & UPH or h$^{-1}$ & 100\% & +50,-50 & 100\% & +50,-50 & 100\% & +50,-50 & Datasheets \\
        \hline
            Wafer Diameter & mm & 100 & +0,-50 & 150 & +33.3,-33.3 & 200 & +0,-25 & I and \cref{fig:wafers} \\
        \hline
            Edge Exclusion & mm & 7 & +0,-50 & 5 & +40,-0 & 5 & +40,-20 & \cite{ledsmagexclusion}\cite{rubiconexclusion} \newline \cite{xiamenexclusion}\cite{american2007annual} \\
        \hline
            Cutting Width & $\mu$m & 100 & +50,-25 & 75 & +33.3,-20 & 20 & +300,-50 & \cite{masaki2000division}\cite{ils2005width} \newline \cite{photonics2010width}\cite{discowidth} \\
        \hline
            Metal Prices & \text{USD}/kg & 100\% & +50,-0 & 100\% & +50,-0 & 100\% & +50,-0 & Datasheets \\
        \hline
            Electricity Price & \text{USD}/kWh & 100\% & +50,-50 & 100\% & +50,-50 & 100\% & +50,-50 & \cite{eia2000electric}\cite{eia2019electric} \\
        \hline
            Saph. Subst. Price & \text{USD} & 40 & +12.5,-12.5 & 10 & +100,-20 & 3 & +66.6,-33.3 & \cref{fig:wafers} \\
        \hline
            Phosphor Prices & \text{USD}/g & 100\% & +50,-50 & 150 & +0,-0 & 200 & +0,-0 & I, \cite{yole_phosphor_2012}\cite{yole2017phosphor} \\
        \hline
        \end{tabularx}
    \vspace{5mm}

@@

\begin{figure}[h!]
 \centering
 \includegraphics[width=15.5cm]{2_SSL_EST/article/figures/subefficiencies_progress.pdf}
 \caption{Historical improvements in light-emitting diode technology. Panels A1-A2 show they key metrics used to compute forward voltage efficiency $\eta_{Vf}$ and droop efficiency $\eta_{droop}$. Panels B1-B2 show the sub-efficiencies internal quantum efficiency $\eta_{IQ}$ and light extraction efficiency $\eta_{LE}$. Source: own synthesis of data from the full list of sources provided in \cref{sec:sources}. Data on average device performance adapted from U.S. Department of Energy (DOE) Reports \cite{doe_ssl_multiyear_2007}\cite{doe_ssl_multiyear_2008}\cite{doe_ssl_multiyear_2013}\cite{doe_ssl_rnd_2016}\cite{doe_ssl_rnd_2018} and an Osram company report \cite{beale_leds_2015}.}
 \label{fgr:subeff}
\end{figure}

@@

\begin{table}[h!]
    \caption{LED innovations and technology improvements affecting key white LED device sub-efficiencies. The table includes innovations and technology improvements affecting forward voltage efficiency, light extraction efficiency, internal quantum efficiency, and light conversion efficiency. For a list of innovations affecting spectral efficiency, see Table 1 in the main article.}
    \vspace{5mm}
    \begin{tabularx}{\textwidth}{ |l|X|X|l|l| }
        \multicolumn{5}{c}{Forward Voltage Efficiency (VFE) $\eta_{Vf}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of Improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            1999 & Indium Tin Oxide \newline Current spreading layer & Contact resistance & Yes & \cite{margalith1999indium}\\
        \hline
            1998 & Digitated electrodes & Contact resistance & No & \cite{steigerwald2001electrode} \\
        \hline
            Ongoing & Epitaxy improvements \newline and better doping & Polarization and \newline bulk resistance & No & I \\
        \hline
            1998 & Silver p-Contacts & Contact resistance & No & \cite{kondoh2001nitride} \\
        \hline
        \multicolumn{5}{c}{Light extraction efficiency (LEE) $\eta_{LE}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            $<$ 2003 & Optimization for \newline cavity effects & Reduces self-interference of quantum well & No & I, \cite{Shen2003} \\
        \hline
            1993 & Chip surface randomization & Total reflection and absorption & No & I, \cite{bergh1973surface}\cite{Schnitzer1993} \\
        \hline
            1993 & Thin-film chip architecture & Absorption & No & I, \cite{Schnitzer1993} \\
        \hline
            1996 & Patterned sapphire substrate  & Total reflection and absorption & Yes & I, \cite{Tadatomo2001} \\
        \hline
            Ongoing & Chip design for high LEE & Total reflection and absorption & No & I, \cite{Haerle2004} \\
        \hline
            $\sim$2000 & Silver p-contacts & Absorption & No & I, \cite{kondoh2001nitride} \\
        \hline
        \multicolumn{5}{c}{Internal quantum efficiency (IQE) $\eta_{IQ}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            1994 & Double heterostructure & Higher Radiative recombination probability & No & \cite{Nakamura1994} \\
        \hline
            1996 & Multiple quantum well & Higher Radiative recombination probability & No & \cite{Koike1996} \\
        \hline
            Ongoing & Active region \newline Doping & Radiative recombination probability & No & I, \cite{schubert2018light} \\
        \hline
            Ongoing & Epitaxy \newline Improvements & Radiative recombination probability & No & I \\
        \hline
            Ongoing & Chip architecture \newline Improvements & Radiative recombination probability & No & I \\
        \hline
        \multicolumn{5}{c}{Light Conversion efficiency (CE) $\eta_{C}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            Ongoing & Lower current density & Current density & No & I \\
        \hline
            Ongoing & Epitaxy \newline Improvements & Charge distribution \newline Defect density & No & I, \cite{bhardwaj2016progress} \\
        \hline
            Ongoing & Chip architecture \newline improvements  & Charge distribution \newline Defect density & No & I, \cite{Wildeson2017} \\
        \hline
            $<$ 2017 & Defect getting \newline underlayer & Defect Density & No & I, \cite{haller2017burying} \\
        \hline
        \end{tabularx}
    \vspace{5mm}

@@

\begin{figure}[H]
	\centering
    \includegraphics[width=0.9\textwidth]{2_SSL_EST/article/figures/phosphor_spectrum-comparison.pdf}
	\caption{Spectral data and additional consumer experience metrics for the earliest identified representative white LED products with published spectral data, shown in Figure 6 in the main article, that used phosphor innovations listed in Table 1 in the main article, each indicated by the phosphor label and spectral data publication year. Top panel: Luminous efficacy of radiation (LER) and colour rendering index (CRI) of white LED devices represented in Figure 6 in the main article. The desirable direction of improvements towards higher luminous efficacy at higher CRI is indicated by a red arrow. Metrics were calculated from spectral data shown in the two bottom panels using the \texttt{colour-science} package for Python \cite{colour-science_software}. Bottom two rows of panels: Corresponding spectral data. The luminosity function \cite{cie-term-lumeff}, describing the wavelength-dependent sensitivity of the human eye, is shown for reference in red in each panel. Note that peaks or large tails of the device emission spectrum at the far ends of the luminosity function are not desirable, as the photons of the corresponding energy are lost to the human eye and count towards the spectral loss channel. Plot legends indicate the years of publication of the spectral data and phosphor mixtures used in corresponding LED devices, with the following designations for additional phosphor mix components: ? -  other parts of phosphor mixture not disclosed, G$^*$ - \ch{CaSrS:Eu^{2+}}, Y$^\star$ - $\beta-$SiAlON:Eu$^{2+}$, G$^\dagger$ - \ch{Lu_3Al_5O_{12}:Ce^{3+}}, O$^\ddagger$ - \ch{(Ba,Sr)_2Si_5N_8:Eu^{2+}}. Source (top panel): own elaboration based on spectral data. Arb. Units - Arbitrary Units, defined as the ratio of radiation intensity at a wavelength compared to the highest point in the spectrum. Sources: top panel - own elaboration based on spectral data; bottom two panels: adapted from published spectral data for LEDs with the following phosphors: YAG \cite{bando1998development}, YGAG \cite{Mueller2002}, 258 \cite{MuellerMach2005}, SLA \cite{Pust2014}, PFS \cite{trigain_spectrum}, QD \cite{lumileds2016qd}\cite{osram2019qd}, SALON \cite{Hoerder2019}.}
\label{fig:phosphor_spectrum}
\end{figure}

@@

\begin{figure}[h]
	\centering
    \includegraphics[width=\textwidth]{2_SSL_EST/article/figures/costmodel_calibration.pdf}
	\caption{Comparison of LED manufacturing cost structure between our manufacturing cost model (bottom panel) and previously published US DOE manufacturing cost calculations and projections based on industry data and LEDCOM model (top panel). Hatched bars are projections given by US DOE for 2018 and 2020. Sources for top panel: \cite{doe2010solid}\cite{doe2011solid}\cite{doe2012solid}\cite{doe2013solid}\cite{doe2014solid}\cite{doe2015solid}\cite{doe2016solid}. Sources for bottom panel: own elaboration based on the cost model described in \cref{sec:costmodel}.}
	\label{fig:costmodel_calibration}
\end{figure}

@@

\begin{table}[h!]
    \begin{tabularx}{\textwidth}{|l|X|}
    \hline
    \textit{Technology} & \textit{References} \\
    \hline
    LED & own research, compare \cite{zenodo_weinold_led_history} \\
    \hline
    CFL (<1984) & \cite{Bouwknegt1982}\cite{Vrenken1983} \\
    \hline
    CFL (1984-2011) & \cite{eger2018origin} \\
    \hline
    CFL (>2011) & \cite{Guan2015} \\
    \hline
    Fire, Incandescent, HID & \cite{azevedo2009transition} augmented by own calculations based on \cite{benesch1905beleuchtungswesen} \\
    \hline
    Max. efficacy & \cite{Murphy2012} \\
    \hline
    \end{tabularx}
\end{table}

@@

\begin{table}[h!]
    \begin{tabularx}{\textwidth}{|l|X|}
    \hline
    \textit{Provider of References} & \textit{References} \\
    \hline
    Stiftung Warentest (Germany) & \cite{Warentest2008}\cite{Warentest2009_1}\cite{Warentest2009_2}\cite{Warentest2010_1}\cite{Warentest2010_2}\cite{Warentest2011}\cite{Warentest2012}\cite{Warentest2013}\cite{Warentest2014_1}\cite{Warentest2014_2}\cite{Warentest2015}\cite{Warentest2016_1}\cite{Warentest2016_2}\cite{Warentest2018} \\
    \hline
    Konsument (Austria) & \cite{Konsument2010} \\
    \hline
    Which (UK) & \cite{Which2020} \\
    \hline
    Industry Periodical & \cite{PM2020} \\
    \hline
    Government Report & \cite{council2013assessment} \\
    \hline
    \end{tabularx}
\end{table}

@@

\begin{table}[h!]
    \begin{tabularx}{\textwidth}{|l|X|}
    \hline
    \textit{Type of Source} & \textit{References} \\
    \hline
    Scientific Publications & \cite{plossl2010wafer}\cite{bierhuizen2007performance}\cite{gencc2019distributed}\cite{chong2014performance} \\
    \hline
    Patents & \cite{patent1999uemura}\cite{patent1998takaoka}\cite{patent1999komaki}\cite{patent1999komaki}\cite{ludowise2006resonant}\cite{camras2005iii}\cite{steigerwald2004contacting} \\
    \hline
    Other Publications & \cite{craford2015}\cite{sun2016}\cite{yole2013packaging} \\
    \hline
    \end{tabularx}
\end{table}

>>>>

\documentclass[parskip=full]{article}

% functionality
\input{auxiliary/preamble/1_preamble_functionality_main}

% layout
\input{auxiliary/preamble/2_preamble_layout_report}

% bibliography
\input{auxiliary/preamble/1_preamble_functionality_bibliography}
\addbibresource{bibliography.bib}

%% CUSTOM SETUP %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% custom figure and table numbering
\renewcommand{\thefigure}{SI\arabic{figure}}
\renewcommand{\thetable}{SI\arabic{table}}

@@

%% METADATA %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\hypersetup{
    pdftitle={Hybrid Life-Cycle Methods},
    pdfauthor={Michael Weinold, Sergey Kolesnikov, Laura Diaz Anadon},
}

%% MAIN DOCUMENT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{figure}[]
	\centering
    \includegraphics[width=\textwidth]{./figures/costmodel_sensitivity.pdf}
	\caption{Sensitivity analysis for selected cost model parameters. Panel a represents the tested range of variation of parameters indicated on horizontal axis along with their respective measurement units, as represented in \cref{tab:sensitivity}. Panel b shows the resulting range of variation in the manufacturing cost. Ranges are represented for three years considered in the cost model, where corresponding data is available, by whiskers of the following colour: 2003 – blue, 2012 – green, 2020 – red. Note the overall trend of decreasing sensitivity to cost model parameters as the number of die per wafer increases over time. Abbreviations: resp. - respective; yield sav. from insp. - yield savings from inspection.}
	\label{fig:sensitivity}
\end{figure}

@@

\begin{table}[]
\small
    \caption{Cost model sensitivity analysis parameter list. The results of the sensitivity analysis for the parameters in this table are presented in \cref{fig:sensitivity}.}
    \vspace{5mm}
    \begin{NiceTabularX}{\textwidth}{ |X|X|l|l|l|l|l|l|X|}
        \hline
            \textit{Parameter} & \textit{Unit} & \textit{2003} & $\pm [\%]$ & \textit{2012} & $\pm [\%]$ & \textit{2020} & $\pm [\%]$ & Source \\
        \hline
            Cleanroom Cost & \text{USD}/m$^2$ & 3000 & +16,-16 & 3000 & +16,-16 & 3000 & +16,-16 & \cite{mddi1997cleanroom}\cite{ledcomv2} \newline \cite{bakshi2009euv}\cite{gajera2006process} \\
        \hline
            Person-hour & FTE & 100\% & +50,-50 & 100\% & +50,-50 & 100\% & +50,-50 & I \\
        \hline
            Equip. Discount & \% of \text{USD} & 0\% & +25,-25 & 0\% & +25,-25 & 0\% & +25,-25 & I \newline \cite{Appleyard_2001} \\
        \hline
            Overall Yield & \% & 100\% & +25,-25 & 100\% & +25,-15 & 100\% & +25,-25 & I, \cite{lumi2012yield}\cite{ledsmag2012} \newline \cite{systemplus2015reverse}\cite{ledcomv2} \\
        \hline
            Inspec. Yield Savings & \%/inspec. & 0.5\% & +80,-80 & 0.5\% & +80,-80 & 0.5\% & +80,-80 & \cite{mckinseyyield} \\
        \hline
            Overall Throughput & UPH or h$^{-1}$ & 100\% & +50,-50 & 100\% & +50,-50 & 100\% & +50,-50 & Datasheets \\
        \hline
            Wafer Diameter & mm & 100 & +0,-50 & 150 & +33.3,-33.3 & 200 & +0,-25 & I and \cref{fig:wafers} \\
        \hline
            Edge Exclusion & mm & 7 & +0,-50 & 5 & +40,-0 & 5 & +40,-20 & \cite{ledsmagexclusion}\cite{rubiconexclusion} \newline \cite{xiamenexclusion}\cite{american2007annual} \\
        \hline
            Cutting Width & $\mu$m & 100 & +50,-25 & 75 & +33.3,-20 & 20 & +300,-50 & \cite{masaki2000division}\cite{ils2005width} \newline \cite{photonics2010width}\cite{discowidth} \\
        \hline
            Metal Prices & \text{USD}/kg & 100\% & +50,-0 & 100\% & +50,-0 & 100\% & +50,-0 & Datasheets \\
        \hline
            Electricity Price & \text{USD}/kWh & 100\% & +50,-50 & 100\% & +50,-50 & 100\% & +50,-50 & \cite{eia2000electric}\cite{eia2019electric} \\
        \hline
            Saph. Subst. Price & \text{USD} & 40 & +12.5,-12.5 & 10 & +100,-20 & 3 & +66.6,-33.3 & \cref{fig:wafers} \\
        \hline
            Phosphor Prices & \text{USD}/g & 100\% & +50,-50 & 150 & +0,-0 & 200 & +0,-0 & I, \cite{yole_phosphor_2012}\cite{yole2017phosphor} \\
        \hline
        \end{NiceTabularX}
    \vspace{5mm}

@@

\begin{figure}[h!]
 \centering
 \includegraphics[width=15.5cm]{./figures/subefficiencies_progress.pdf}
 \caption{Historical improvements in light-emitting diode technology. Panels A1-A2 show they key metrics used to compute forward voltage efficiency $\eta_{Vf}$ and droop efficiency $\eta_{droop}$. Panels B1-B2 show the sub-efficiencies internal quantum efficiency $\eta_{IQ}$ and light extraction efficiency $\eta_{LE}$. Source: own synthesis of data from the full list of sources provided in \cref{sec:sources}. Data on average device performance adapted from U.S. Department of Energy (DOE) Reports \cite{doe_ssl_multiyear_2007}\cite{doe_ssl_multiyear_2008}\cite{doe_ssl_multiyear_2013}\cite{doe_ssl_rnd_2016}\cite{doe_ssl_rnd_2018} and an Osram company report \cite{beale_leds_2015}.}
 \label{fgr:subeff}
\end{figure}

@@

\begin{table}[h!]
    \caption{LED innovations and technology improvements affecting key white LED device sub-efficiencies. The table includes innovations and technology improvements affecting forward voltage efficiency, light extraction efficiency, internal quantum efficiency, and light conversion efficiency. For a list of innovations affecting spectral efficiency, see Table 1 in the main article.}
    \vspace{5mm}
    \begin{NiceTabularX}{\textwidth}{ |l|X|X|l|l| }
        \multicolumn{5}{c}{Forward Voltage Efficiency (VFE) $\eta_{Vf}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of Improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            1999 & Indium Tin Oxide \newline Current spreading layer & Contact resistance & Yes & \cite{margalith1999indium}\\
        \hline
            1998 & Digitated electrodes & Contact resistance & No & \cite{steigerwald2001electrode} \\
        \hline
            Ongoing & Epitaxy improvements \newline and better doping & Polarization and \newline bulk resistance & No & I \\
        \hline
            1998 & Silver p-Contacts & Contact resistance & No & \cite{kondoh2001nitride} \\
        \hline
        \multicolumn{5}{c}{Light extraction efficiency (LEE) $\eta_{LE}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            $<$ 2003 & Optimization for \newline cavity effects & Reduces self-interference of quantum well & No & I, \cite{Shen2003} \\
        \hline
            1993 & Chip surface randomization & Total reflection and absorption & No & I, \cite{bergh1973surface}\cite{Schnitzer1993} \\
        \hline
            1993 & Thin-film chip architecture & Absorption & No & I, \cite{Schnitzer1993} \\
        \hline
            1996 & Patterned sapphire substrate  & Total reflection and absorption & Yes & I, \cite{Tadatomo2001} \\
        \hline
            Ongoing & Chip design for high LEE & Total reflection and absorption & No & I, \cite{Haerle2004} \\
        \hline
            $\sim$2000 & Silver p-contacts & Absorption & No & I, \cite{kondoh2001nitride} \\
        \hline
        \multicolumn{5}{c}{Internal quantum efficiency (IQE) $\eta_{IQ}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            1994 & Double heterostructure & Higher Radiative recombination probability & No & \cite{Nakamura1994} \\
        \hline
            1996 & Multiple quantum well & Higher Radiative recombination probability & No & \cite{Koike1996} \\
        \hline
            Ongoing & Active region \newline Doping & Radiative recombination probability & No & I, \cite{schubert2018light} \\
        \hline
            Ongoing & Epitaxy \newline Improvements & Radiative recombination probability & No & I \\
        \hline
            Ongoing & Chip architecture \newline Improvements & Radiative recombination probability & No & I \\
        \hline
        \multicolumn{5}{c}{Light Conversion efficiency (CE) $\eta_{C}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            Ongoing & Lower current density & Current density & No & I \\
        \hline
            Ongoing & Epitaxy \newline Improvements & Charge distribution \newline Defect density & No & I, \cite{bhardwaj2016progress} \\
        \hline
            Ongoing & Chip architecture \newline improvements  & Charge distribution \newline Defect density & No & I, \cite{Wildeson2017} \\
        \hline
            $<$ 2017 & Defect getting \newline underlayer & Defect Density & No & I, \cite{haller2017burying} \\
        \hline
        \end{NiceTabularX}
    \vspace{5mm}

@@

\begin{figure}[H]
	\centering
    \includegraphics[width=0.9\textwidth]{./figures/phosphor_spectrum-comparison.pdf}
	\caption{Spectral data and additional consumer experience metrics for the earliest identified representative white LED products with published spectral data, shown in Figure 6 in the main article, that used phosphor innovations listed in Table 1 in the main article, each indicated by the phosphor label and spectral data publication year. Top panel: Luminous efficacy of radiation (LER) and colour rendering index (CRI) of white LED devices represented in Figure 6 in the main article. The desirable direction of improvements towards higher luminous efficacy at higher CRI is indicated by a red arrow. Metrics were calculated from spectral data shown in the two bottom panels using the \texttt{colour-science} package for Python \cite{colour-science_software}. Bottom two rows of panels: Corresponding spectral data. The luminosity function \cite{cie-term-lumeff}, describing the wavelength-dependent sensitivity of the human eye, is shown for reference in red in each panel. Note that peaks or large tails of the device emission spectrum at the far ends of the luminosity function are not desirable, as the photons of the corresponding energy are lost to the human eye and count towards the spectral loss channel. Plot legends indicate the years of publication of the spectral data and phosphor mixtures used in corresponding LED devices, with the following designations for additional phosphor mix components: ? -  other parts of phosphor mixture not disclosed, G$^*$ - \ch{CaSrS:Eu^{2+}}, Y$^\star$ - $\beta-$SiAlON:Eu$^{2+}$, G$^\dagger$ - \ch{Lu_3Al_5O_{12}:Ce^{3+}}, O$^\ddagger$ - \ch{(Ba,Sr)_2Si_5N_8:Eu^{2+}}. Source (top panel): own elaboration based on spectral data. Arb. Units - Arbitrary Units, defined as the ratio of radiation intensity at a wavelength compared to the highest point in the spectrum. Sources: top panel - own elaboration based on spectral data; bottom two panels: adapted from published spectral data for LEDs with the following phosphors: YAG \cite{bando1998development}, YGAG \cite{Mueller2002}, 258 \cite{MuellerMach2005}, SLA \cite{Pust2014}, PFS \cite{trigain_spectrum}, QD \cite{lumileds2016qd}\cite{osram2019qd}, SALON \cite{Hoerder2019}.}
\label{fig:phosphor_spectrum}
\end{figure}

@@

\begin{figure}[h]
	\centering
    \includegraphics[width=\textwidth]{./figures/costmodel_calibration.pdf}
	\caption{Comparison of LED manufacturing cost structure between our manufacturing cost model (bottom panel) and previously published US DOE manufacturing cost calculations and projections based on industry data and LEDCOM model (top panel). Hatched bars are projections given by US DOE for 2018 and 2020. Sources for top panel: \cite{doe2010solid}\cite{doe2011solid}\cite{doe2012solid}\cite{doe2013solid}\cite{doe2014solid}\cite{doe2015solid}\cite{doe2016solid}. Sources for bottom panel: own elaboration based on the cost model described in \cref{sec:costmodel}.}
	\label{fig:costmodel_calibration}
\end{figure}

@@

\begin{table}[h!]
    \begin{NiceTabularX}{\textwidth}{|l|X|}
    \hline
    \textit{Technology} & \textit{References} \\
    \hline
    LED & own research, compare \cite{zenodo_weinold_led_history} \\
    \hline
    CFL (<1984) & \cite{Bouwknegt1982}\cite{Vrenken1983} \\
    \hline
    CFL (1984-2011) & \cite{eger2018origin} \\
    \hline
    CFL (>2011) & \cite{Guan2015} \\
    \hline
    Fire, Incandescent, HID & \cite{azevedo2009transition} augmented by own calculations based on \cite{benesch1905beleuchtungswesen} \\
    \hline
    Max. efficacy & \cite{Murphy2012} \\
    \hline
    \end{NiceTabularX}
\end{table}

@@

\begin{table}[h!]
    \begin{NiceTabularX}{\textwidth}{|l|X|}
    \hline
    \textit{Provider of References} & \textit{References} \\
    \hline
    Stiftung Warentest (Germany) & \cite{Warentest2008}\cite{Warentest2009_1}\cite{Warentest2009_2}\cite{Warentest2010_1}\cite{Warentest2010_2}\cite{Warentest2011}\cite{Warentest2012}\cite{Warentest2013}\cite{Warentest2014_1}\cite{Warentest2014_2}\cite{Warentest2015}\cite{Warentest2016_1}\cite{Warentest2016_2}\cite{Warentest2018} \\
    \hline
    Konsument (Austria) & \cite{Konsument2010} \\
    \hline
    Which (UK) & \cite{Which2020} \\
    \hline
    Industry Periodical & \cite{PM2020} \\
    \hline
    Government Report & \cite{council2013assessment} \\
    \hline
    \end{NiceTabularX}
\end{table}

@@

\begin{table}[h!]
    \begin{NiceTabularX}{\textwidth}{|l|X|}
    \hline
    \textit{Type of Source} & \textit{References} \\
    \hline
    Scientific Publications & \cite{plossl2010wafer}\cite{bierhuizen2007performance}\cite{gencc2019distributed}\cite{chong2014performance} \\
    \hline
    Patents & \cite{patent1999uemura}\cite{patent1998takaoka}\cite{patent1999komaki}\cite{patent1999komaki}\cite{ludowise2006resonant}\cite{camras2005iii}\cite{steigerwald2004contacting} \\
    \hline
    Other Publications & \cite{craford2015}\cite{sun2016}\cite{yole2013packaging} \\
    \hline
    \end{NiceTabularX}
\end{table}
